\chapter{Coordinates with Mutable Successors}

\begin{lemma}
	For each of the ranges $[h, t]$ returned by \verb|split_linked|, $h = l \Leftrightarrow t = l$.
\end{lemma}

\begin{exercise}
	Assuming that one of the ranges $(h, t)$ returned by \verb|split_linked| is not empty, explain what
	iterator $t$ points to and what the value of $\func{successor}(t)$ is.
\end{exercise}

\begin{solution}
	If one of the ranges $(h, t)$ is not empty, then $t$ points to the last iterator in the range $[f, l)$ such
	that $\forall i \in [successor(t), l) p(i) \neq p(t)$. For the successor we have either
	\[ \func{successor}(t) = l \vee p(t) \neq p(\func{successor}(t)). \]
\end{solution}

\begin{lemma}
	\verb|split_linked| is a precedence-preserving link rearrangement.
\end{lemma}

\begin{proof}
	\verb|split_linked| always links the tail to an iterator that succeeds itself once the corresponding
	head has been assigned.
\end{proof}

\begin{exercise}
	Implement \verb|combine_linked_nonempty|, allowing for empty inputs. What value should be returned as the
	last-visited iterator?
\end{exercise}

\lstinputlisting[language=c++, firstline=3264, lastline=3276]{eop-code/eop.h}

\begin{lemma}
	If a call $\func{combine\_linked\_nonempty}(f0, l0, l1, r, s)$ returns $(h, t, l)$, $h$ equals $f0$ or $f1$,
	and, independently, $l$ equals $l0$ or $l1$.
\end{lemma}

\begin{proof}
	At the start of the function, $h$ is either set to $f0$ or $f1$ and is not set again. The function
	has only two exit points and in one case $l$ equals $l0$ and in the other case $l$ equals $l1$.
\end{proof}

\begin{lemma}
	When state $s2$ is reached, $t$ is from the original range $[f0, l0)$, $\func{successor}(t) = l0$, and
	$f1 \neq l1$; when state $s3$ is reached, $t$ is from the original range $[f1, l1)$,
	$\func{successor}(t) = l1$, and $f0 \neq l0$.
\end{lemma}

\begin{proof}
	State $s2$ is only reached from state $s0$ when $f0 = l0$. There are three cases where we reach state
	$s0$. In two of the cases we call $\func{advance\_tail}(t, f0)$ before going to state $s0$ and in the other
	case we call $\func{link\_to\_tail}(t, f0)$. In each case $t$ points to $f0$ and then $f0$ is advanced.
	Since we just assigned $t$ to $f0$ before advancing we have $\func{successor}(t) = l0$.
\end{proof}

\begin{lemma}
	\verb|combine_linked_nonempty| is a precedence-preserving link rearrangement.
\end{lemma}

\begin{proof}
	\verb|combine_linked_nonempty| always links the tail to an iterator that succeeds itself once the
	corresponding head has been assigned.
\end{proof}

\begin{lemma}
	If $[f0, l0)$ and $[f1, l1)$ are nonempty increasing bounded ranges, their merge with
	\verb|merge_linked_nonempty| is an increasing bounded range.
\end{lemma}

\begin{proof}
	Since \verb|combine_linked_nonempty| is a precedence-preserving link rearrangement (Lemma 8.5), and
	our input ranges are increasing bounded ranges, \verb|merge_linked_nonempty| is an increasing bounded
	range.
\end{proof}

\begin{lemma}
	If $i0 \in [f0, l0)$ and $i1 \in [f1, l1)$ are iterators whose values are equivalent under $r$, in the
	merge of these ranges with \verb|merge_linked_nonempty|, $i0 \prec i1$.
\end{lemma}

\begin{proof}
	\verb|combine_linked_nonempty| is a precedence-preserving link rearrangement (Lemma 8.5).
\end{proof}

\begin{lemma}
	\verb|sort_linked_nonempty_n| is a link rearrangement.
\end{lemma}

\begin{proof}
	\verb|merge_linked_nonempty| is a link rearrangement since \verb|combine_linked_nonempty| is a link
	rearrangement.
\end{proof}

\begin{lemma}
	If $\llbracket f, n \rrparenthesis$ is a nonempty counted range, \verb|sort_linked_nonempty_n| will
	rearrange it into an increasing bounded range.
\end{lemma}

\begin{lemma}
	\verb|sort_linked_nonempty_n| is stable with respect to the supplied weak ordering $r$.
\end{lemma}

\begin{exercise}
	Determine the formulas for the worst-case and average number of applications of the relation and of the
	linker object in \verb|sort_linked_nonempty_n|.
\end{exercise}

\begin{exercise}
	Implement a precedence-preserving linked rearrangement \verb|unique| that takes a linked range and an
	equivalence relation on the value type of the iterators and that produces two ranges by moving all except
	the first iterator in any adjacent sequence of iterators with equivalent values to a second range.
\end{exercise}

\lstinputlisting[language=c++, firstline=3382, lastline=3402]{eop-code/eop.h}
